Refinement for open automata

### 2021-05-11 2021-Refinement/report.tex
Added TODOs

The relations will have to be at least preorders, simulations and \Ludo{introduction of deadlocks}{define more precisely introduction of deadlocks or refer to a future definition} (state where no transition is possible) to be called refinement simulation.
On top of that weak refinement are refinement where some actions, typically one named \(\tau\), are considered not observable and can occur in the implementation while not having to be matched in the specification.
In the weak refinement the relations have to forbid the introduction of livelocks (only a sequence of \(\tau\) possible) and respect \(\tau\)-stuttering \marginpar{cite many things arround here}.
\Ludo{to add}{I believe we want more arguments on why we want some properties, at least deadlock and livelock related, perhaps also weak vs strong. How did you decide of the list of useful properties / when ais such or such property useful?}

@@

Two states \(s_1 \in S_1, s_2 \in S_2\) are related under a \Ludo{?}{relation is mission: under a relation $R$ for a valuation ... ?} valuation \(\sigma: V_1 \uplus V_2 \to \values\) iff \(\sigma \vdash R\mpar{s_1, s_2}\).
For a relation \(\wrel{}{}{}\) which uses a relation under predicate to prove that two element are related \(\forall x \, x', x \wrel{}{}{} x' \iff \exists R, P\mpar{R}\), the fact that there is a valid witness \(R\) is noted \(\wrel{x}{x'}{R}\).
\end{defi}
Now the properties about states of automata can be adapted to open automata.
\begin{defi}[Simulation on open automata]
A relation \(\leq\) is a simulation if for any two related automata \(\wrel{A_1}{A_2}{R}\) where \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\), \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\) and \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\), \(R\) satisfies both
\defitem Initial states are related under initial valuations: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\);
\defitem From related states, all out-transitions from \(A_1\) can be simulated in \(A_2\) and their target states are \Ludo{related}{I am a bit surprised that $J_1$ and $J_2$ are unrelated and we only need to take only the intersection TBD}:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \forall \sigma: V_1 \uplus V_2 \uplus \fvars{t_1} \to \values, \\
	\mpar{\sigma \vdash R\mpar{s_1, s_2} \wedge g_1} \implies \bigsymb{\exists} t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge g_2 \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
The last formula means that for every pair of related states under valuation of the automaton variables and every possible transitions (\(\vdash g_1\)) from the first automaton, there is a possible transition (\(\vdash g_2\)) such that the produced action matches (\(\alpha_1 = \alpha_2\)), the holes actions from same hole names (\(J'_1 \cap J'_2\)) match (\(\beta_{1j} = \beta_{2j}\)) and the target states are related after variable update.
This is a natural extension of the notion of simulation on LTS, which is the same definition without variables and holes and guards.
It will serve as a watchdog for any relation that will be defined with ``simulation" in its name, except weak-simulation where the transition from the first automaton that produce \(\tau\) actions are excluded.

@@

``Deadlock reduction" conflicts with ``composition completeness" by disallowing some unwanted cases where an automaton in a hole cannot produce any action that any out-transitions expects.
In particular filling a hole with the deadlock automaton (only one state without out-transitions) is not considered a refinement for a \Ludo{no deadlock reducing simulation refinement}{define or use the name of the definition}.
A way to solve this \Ludo{issue}{why is this an issue? can be solved by stating which properties are desirable for which reason above} is to characterise a composition that do not introduce deadlocks, which will be used instead of the composition introduced earlier.
\begin{defi}[Reachability]
For any open automata \(A \defobject \OA{S}{s_0}{V}{\sigma_0}{J}{T}\), a reachability predicate \(\reach{A}: S \to \rformulae[V]\) is a predicate on states such that
\defitem \(\sigma_0 \vdash \reach{A}\mpar{s_0}\).
\defitem \nmm{\bigsymb{\forall} t \defobject \OTg \in T, \forall \nu: \fvars{t} \to \values, \nu \vdash \reach{A}\mpar{s} \wedge g \implies \reach{A}\mpar{s'}\psubst{\psi}}.
\end{defi}

@@

The composition  \(A_p\subst{A_c}{k} = \OA{S}{s_0}{V}{\sigma_0}{J}{T}\) with \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\), \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) and \(k \in J_p\), is a non-locking composition if there is a reachability predicate such that from reachable configurations, either a transition is possible in \(A_p\subst{A_c}{k}\) or there were no possible transition in \(A_p\) to begin with
\begin{multline*}
	\forall s \in S, \forall \sigma: V \to \values, \mpar{\sigma \vdash \reach{A_p\subst{A_c}{k}}\mpar{s}} \implies \mpar{\exists t \in \fOT{s}, \exists \nu: \fvars{t} \to \values, \sigma \uplus \nu \vdash \fguard{t}} \\
	\vee \forall t_p \in \fOT{\pi_{S_p}\mpar{s}}, \forall \nu: \fvars{t_p} \to \values, \sigma \uplus \nu \nvdash \fguard{t_p}
\end{multline*}
\end{defi}
(\(\pi_X\) is the projection of tuples on \(X\)).

@@

When expanding the aliases \(S, i, V\) in \Ludo{the definition}{which def? pas lu la suite car je savais pas ou ca allait}, and remarking that the last existential quantifier can be unfolded into a transition in \(A\) and one in \(A_p\) with the added condition that they match, the definition become mildly similar with the deadlock reduction one.
It is not a coincidence because their goal is to caracterise the same kind of compatibility between automata.
Actually it is possible to simplify both definition and make them even more similar.

\begin{defi}[Deadlock reduction, working definition]
A simulation \(\leq\) is deadlock reducing if in all the \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\) such that \(\wrel{A_1}{A_2}{R}\), there is one that also satisfies
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V_1 \uplus V_2 \uplus \bigcup_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \to \values, \\
	\sigma \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}
\end{multline*}
\Ludo{New Question}{what is the meaning of sigma here: it is not used on the right hand side of the implication; this seem to mean if there is  a sigma such that s1 and s2 are related then we have thr peoperty on disujunction on pred qithout caring about the valuation. Am I right? same question for the def below}
\end{defi}
\begin{lem}
The intuitive and working definition of deadlock reduction are equivalent.
\end{lem}

@@

\section{Preliminary refinement relations}\label{sec:prelref}
\Ludo{organisation}{Is this only preliminaries, if yes could it be placed earlier? if no perhaps rename into refinement for open automata something}
The main goal of this section is to help the reader understand the problems, constraints and different elements involved to solve them so that they are not all introduced at once in the real refinement relation.
The main objective of the refinement relations \Ludo{in this section}{du coup la on se pose la question est ce juste ici ou est ce le but global, c est quio le schema general? ceci dit ca sera peut etre resolu par l introduction ... mais dans tous les cas une petite couche sur ou on est et ou on va ca serait bien. il faut quelque part clairemetn definir ce qu on veut et sous quelle condition plus tot que ici} is to be able to say that a composition of two automata is a refinement of the base automaton (\(a\mbrk{b} \leq a\)) in restricted interesting cases.
\Ludo{lost}{I do not understand the last sentence}
Hopefully they can be used as simpler (and less expensive) versions of the real refinement relation in the case the manipulated automata respect the constraints.

@@

For two open automata \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), we first define a refinement relation where holes of the compared automata are identical.
This relation is the basis for all the other ones.
The goal is to characterise open automata that have a more determined (more deterministic) behaviour without adding deadlocks. This corresponds to ???? in the classical definitions of refinement \\Ludo{TODO}{j imagine que rabea peut donner des refs ici}
\begin{defi}[Hole-identical refinement] % TODO: transition variables
If \(J_1 = J_2\) then ``\(A_1\) is a hole-identical refinement of \(A_2\)", noted \(A_1 \leq_= A_2\), is defined as:
\[ \exists R: \mpar{S_1 \times S_2} \to \rformulae[V_1 \uplus V_2], \sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}} \qwedge R \text{ hole-identical simulation of } A_1 \text{ in } A_2 \]
\end{defi}
As in other simulation-like relations, \(R\) is a witness of \(A_1 \leq_= A_2\).
The standard requirements for (bi)simulations are that \(R\) relates initial states, transitions can be matched, and target of matched transitions with related sources are also related.
For a bisimulation transitions are matched one-to-many from each automaton to the other, but for a simulation only one way is used.
The definition above requires explicitely that initial states are related, the other requirement are encapsulated in the definition (below) of hole-identical simulation.

The specificities of open automata at on the relation between the (initial) states.
\Ludo{lost}{When reading the next sentences I am not sure where we are going, what is notation introduction of concepts or definitions, please clarify}
The relation \(R\) relates \(i_1\) to \(i_2\) under the condition that the current value of variables satisfies \(R\mpar{i_1, i_2}\).
If two states \(s_1 \in S_1, s_2 \in S_2\) are not(/never) related then \(\nvdash R\mpar{s_1, s_2}\).
The predicate \(R\mpar{i_1, i_2}\) is used to take into account the fact that a state is also constituted of the value of the variables.
This is already used in FH-bisimulation introduced in previous articles about open automata.
\Ludo{OK}{I guess I understand that you want to reach the next sentence as an explanation of the notations used in the definition, I mainly believe it is necessary to re-organise the ideas before}
So the meaning of \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\) is that the initial states with the initial valuation are effectively related.
\newpage % TODO: remove when less comments :)

\begin{defi}[Hole-identical simulation]
\Quentin{``\(R\) is a hole-identical simulation of \(A_1\) in \(A_2\)"}{Any idea on how to note that? \(R \vDash A_1 \leq_= A_2\) for instance.} \Ludo{reply}{I believe the notation  \(R \vDash A_1 \leq_= A_2\) is not bad bbut for the text I would rather write a1 is a simulation of a2 according to (as witnessed by) R ???} is defined as: % TODO: \wrel
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \\
	\mpar{\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\forall \sigma: \mpar{V_1 \uplus \fvars{t_1} \uplus V_2} \to \values, \\
		\quad \sigma \vdash R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_{j \in J'} \beta_{1j} = \beta_{2xj} \\[12pt]
			\nwedge g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}} \\[12pt]
	\end{array}} \\
	\wedge \forall \sigma: \mpar{V_1 \uplus V_2} \to \values, \sigma \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}
\end{multline*}
\end{defi}
The first two lines quantify on source states, transitions (and target states) and source states valuations.
The \Ludo{3\textsuperscript{rd}}{is this the good line number?} line is the condition to ensure that transitions are not matched to incompatible ones:
For all variable assignments of the two automata and all hole actions (quantification of \(\sigma\)), assuming that the states were related and the transition in \(A_1\) is possible (\(R\mpar{s_1, s_2} \wedge g_1\) part) then there is always a set of transitions, not necessarily only one as several transitions might simulate the same one depending on the value of variables for example; these transitions can be performed (\(g_{2x}\)), produce the same action (\(\alpha_1 = \alpha_{2x}\)), accept the same action from the holes (\(\beta_{1j} = \beta_{2xj}\)) and satisfy the predicate for relating the target states after variable update (\(R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}\)).

>>>>

The relations will have to be at least preorders, simulations and prevent the introduction of deadlocks to be called refinement simulation.
Deadlocks are states and valuations that are reachable in a run and from which no further transition is possible.
On top of that weak refinement are refinement where some actions, typically one named \(\tau\), are considered not observable and can occur in the implementation while not having to be matched in the specification.
In the weak refinement the relations have to forbid the introduction of livelocks (only a sequence of \(\tau\) possible) and respect \(\tau\)-stuttering \marginpar{cite many things arround here}.
\Ludo{to add}{I believe we want more arguments on why we want some properties, at least deadlock and livelock related, perhaps also weak vs strong. How did you decide of the list of useful properties / when ais such or such property useful?}

@@

Two states \(s_1 \in S_1, s_2 \in S_2\) with their respective valuations \(\sigma_1: V_1 \to \values, \sigma_2: V_2 \to \values\) are related iff \(\sigma_1 \uplus \sigma_2 \vdash R\mpar{s_1, s_2}\).
For a relation \(\wrel{}{}{}\) which uses a relation under predicate to prove that two element are related \(\forall x \, x', x \wrel{}{}{} x' \iff \exists R, P\mpar{R}\), the fact that there is a valid witness \(R\) is noted \(\wrel{x}{x'}{R}\).
\end{defi}
Now the properties about states of automata can be adapted to open automata.
\begin{defi}[Simulation on open automata]
A relation \(\leq\) is a simulation if for any two related automata \(\wrel{A_1}{A_2}{R}\) where \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\), \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\) and \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\), \(R\) satisfies both
\defitem Initial states are related under initial valuations: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\);
\defitem From related states, all out-transitions from \(A_1\) can be simulated in \(A_2\) and their target states are \Ludo{related}{I am a bit surprised that $J_1$ and $J_2$ are unrelated and we only need to take only the intersection TBD}:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \forall \sigma: V_1 \uplus V_2 \uplus \fvars{t_1} \to \values, \\
	\mpar{\sigma \vdash R\mpar{s_1, s_2} \wedge g_1} \implies \bigsymb{\exists} t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge g_2 \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
The last formula means that for every pair of related states under valuation of the automaton variables and every possible transitions (\(\vdash g_1\)) from the first automaton, there is a possible transition (\(\vdash g_2\)) such that the produced action matches (\(\alpha_1 = \alpha_2\)), the holes actions from same hole names (\(J'_1 \cap J'_2\)) match (\(\beta_{1j} = \beta_{2j}\)) and the target states are related after variable update.
This is a natural extension of the notion of simulation on LTS, which is the same definition without variables and holes and guards.
It will serve as a watchdog for any relation that will be defined with ``simulation" in its name, except weak-simulation where the transition from the first automaton that produce \(\tau\) actions are excluded.

@@

``Deadlock reduction" conflicts with ``composition completeness" by disallowing some unwanted cases where an automaton in a hole cannot produce any action that any out-transitions expects.
In particular filling a hole with the deadlock automaton (only one state without out-transitions) is not considered a refinement for a \Ludo{deadlock reducing simulation refinement}{define or use the name of the definition}.
A way to solve this \Ludo{issue}{why is this an issue? can be solved by stating which properties are desirable for which reason above} is to characterise a composition that do not introduce deadlocks, which will be used instead of the composition introduced earlier.
\begin{defi}[Reachability]
For any open automata \(A \defobject \OA{S}{s_0}{V}{\sigma_0}{J}{T}\), a reachability predicate \(\reach{A}: S \to \rformulae[V]\) is a predicate on states such that
\defitem \(\sigma_0 \vdash \reach{A}\mpar{s_0}\).
\defitem \nmm{\bigsymb{\forall} t \defobject \OTg \in T, \forall \nu: \fvars{t} \to \values, \nu \vdash \reach{A}\mpar{s} \wedge g \implies \reach{A}\mpar{s'}\psubst{\psi}}.
\end{defi}

@@

The composition  \(A_p\subst{A_c}{k} = \OA{S}{s_0}{V}{\sigma_0}{J}{T}\) with \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\), \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) and \(k \in J_p\), is a non-locking composition if there is a reachability predicate such that from reachable configurations, either a transition is possible in \(A_p\subst{A_c}{k}\) or there were no possible transition in \(A_p\) to begin with
\begin{multline*}
	\forall s \in S, \forall \sigma: V \to \values, \mpar{\sigma \vdash \reach{A_p\subst{A_c}{k}}\mpar{s}} \implies \mpar{\exists t \in \fOT{s}, \exists \nu: \fvars{t} \to \values, \sigma \uplus \nu \vdash \fguard{t}} \\
	\vee \forall t_p \in \fOT{\pi_{S_p}\mpar{s}}, \forall \nu: \fvars{t_p} \to \values, \sigma \uplus \nu \nvdash \fguard{t_p}
\end{multline*}
\end{defi}
(\(\pi_X\) is the projection of tuples on the set \(X\)). % TODO: Move that to the notations

@@

When expanding the aliases \(S, s_0, V\) in \Ludo{the definition}{which def? pas lu la suite car je savais pas ou ca allait}, and remarking that the last existential quantifier can be unfolded into a transition in \(A\) and one in \(A_p\) with the added condition that they match, the definition become mildly similar with the deadlock reduction one. % TODO: Make it clearer, what expansion, what is the result; or remove
It is not a coincidence because their goal is to caracterise the same kind of compatibility between automata.
Actually it is possible to simplify both definition and make them even more similar.

\begin{defi}[Deadlock reduction, working definition]
A simulation \(\leq\) is deadlock reducing if in all the \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\) such that \(\wrel{A_1}{A_2}{R}\), there is one that also satisfies
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V_1 \uplus V_2 \uplus \bigcup_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \to \values, \\
	\sigma \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}
\end{multline*}
\Ludo{New Question}{what is the meaning of sigma here: it is not used on the right hand side of the implication; this seem to mean if there is  a sigma such that s1 and s2 are related then we have thr peoperty on disujunction on pred qithout caring about the valuation. Am I right? same question for the def below}
\Quentin{Reply}{Any variable not defined by the valuation is under an implicit \(\exists\) quantifier (satisfiability). What it means is that for every valuation where states are related and a transition is possible (disjunction on the left), there is a value for the free variables of the disjunction on the right that makes a transition possible (the existential quantifier has been wisely displaced in this explaination to make it easier to understand).}
\end{defi}
% TODO: Explain this formulations (it will also do the job of explaining the definition below)
\begin{lem}
The intuitive and working definition of deadlock reduction are equivalent.
\end{lem}

@@

\section{Refinement relations in restricted cases}\label{sec:prelref}
\Ludo{organisation}{Is this only preliminaries, if yes could it be placed earlier? if no perhaps rename into refinement for open automata something}
The main goal of this section is to help the reader understand the problems, constraints and different elements involved to solve them so that they are not all introduced at once in the real refinement relation.
The main objective of the refinement relations \Ludo{in this section}{du coup la on se pose la question est ce juste ici ou est ce le but global, c est quio le schema general? ceci dit ca sera peut etre resolu par l introduction ... mais dans tous les cas une petite couche sur ou on est et ou on va ca serait bien. il faut quelque part clairemetn definir ce qu on veut et sous quelle condition plus tot que ici} is to be able to state that a composition of two automata is a refinement of the base automaton (\(a\mbrk{b} \leq a\)) in restricted interesting cases. % TODO: probably remove, this is was written when the previous section was not yet written
\Ludo{lost}{I do not understand the last sentence}
Hopefully they can be used as simpler (and less expensive) versions of the real refinement relation in the case the manipulated automata respect the constraints.

@@

For two open automata \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), we first define a refinement relation where holes of the compared automata are identical.
This relation is the basis for all the other ones.
The goal is to characterise open automata that have a more determined (more deterministic) behaviour without adding deadlocks.
This corresponds to ???? in the classical definitions of refinement \Ludo{TODO}{j imagine que rabea peut donner des refs ici}.
\begin{defi}[Hole-identical refinement] % TODO: transition variables
If \(J_1 = J_2\) then ``\(A_1\) is a hole-identical refinement of \(A_2\)", noted \(A_1 \leq_= A_2\), is defined as:
\[ \exists R: \mpar{S_1 \times S_2} \to \rformulae[V_1 \uplus V_2], \sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}} \qwedge R \text{ hole-identical simulation of } A_1 \text{ in } A_2 \]
\end{defi}
As in other simulation-like relations, \(R\) is a witness of \(A_1 \leq_= A_2\).
The standard requirements for (bi)simulations are that \(R\) relates initial states, transitions can be matched, and target of matched transitions with related sources are also related.
For a bisimulation transitions are matched one-to-many from each automaton to the other, but for a simulation only one way is used.
The definition above requires explicitely that initial states are related, the other requirement are encapsulated in the definition (below) of hole-identical simulation.

The specificities of open automata at on the relation between the (initial) states.
\Ludo{lost}{When reading the next sentences I am not sure where we are going, what is notation introduction of concepts or definitions, please clarify}
The relation \(R\) relates \(i_1\) to \(i_2\) under the condition that the current value of variables satisfies \(R\mpar{i_1, i_2}\).
If two states \(s_1 \in S_1, s_2 \in S_2\) are not(/never) related then \(\nvdash R\mpar{s_1, s_2}\).
The predicate \(R\mpar{i_1, i_2}\) is used to take into account the fact that a state is also constituted of the value of the variables.
This is already used in FH-bisimulation introduced in previous articles about open automata.
\Ludo{OK}{I guess I understand that you want to reach the next sentence as an explanation of the notations used in the definition, I mainly believe it is necessary to re-organise the ideas before}
So the meaning of \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\) is that the initial states with the initial valuation are effectively related. % TODO: Also a relic of the past, when things were explained on the go instead of in a dedicated section. Remove if already explained, move otherwise
\newpage % TODO: remove when less comments :)

\begin{defi}[Hole-identical simulation]
\Quentin{``\(R\) is a hole-identical simulation of \(A_1\) in \(A_2\)"}{Any idea on how to note that? \(R \vDash A_1 \leq_= A_2\) for instance.} \Ludo{reply}{I believe the notation  \(R \vDash A_1 \leq_= A_2\) is not bad bbut for the text I would rather write a1 is a simulation of a2 according to (as witnessed by) R ???} is defined as: % TODO: Actually this was already answered, \wrel is the settled way
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \\
	\mpar{\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\forall \sigma: \mpar{V_1 \uplus \fvars{t_1} \uplus V_2} \to \values, \\
		\quad \sigma \vdash R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_{j \in J'} \beta_{1j} = \beta_{2xj} \\[12pt]
			\nwedge g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}} \\[12pt]
	\end{array}} \\
	\wedge \forall \sigma: \mpar{V_1 \uplus V_2} \to \values, \sigma \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}
\end{multline*}
\end{defi}
The first two lines quantify on source states, transitions (and target states) and source states valuations.
The \Ludo{3\textsuperscript{rd}}{is this the good line number?} line is the condition to ensure that transitions are not matched to incompatible ones:
For all variable assignments of the two automata and all hole actions (quantification of \(\sigma\)), assuming that the states were related and the transition in \(A_1\) is possible (\(R\mpar{s_1, s_2} \wedge g_1\) part) then there is always a set of transitions, not necessarily only one as several transitions might simulate the same one depending on the value of variables for example; these transitions can be performed (\(g_{2x}\)), produce the same action (\(\alpha_1 = \alpha_{2x}\)), accept the same action from the holes (\(\beta_{1j} = \beta_{2xj}\)) and satisfy the predicate for relating the target states after variable update (\(R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}\)).
